A White Noise Approach to Evolutionary Ecology

### 2020-04-18 .A White Noise Approach To Evolutionary Ecology/draft.tex
split main text and SM

\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage[left]{lineno}
\usepackage{amsmath}
\linenumbers
\usepackage{float}
\usepackage{flafter}

@@

\hypertarget{supplementary-material-sm}{%
\section{Supplementary material (SM)}\label{supplementary-material-sm}}

Throughout this supplement, we set use dot notation for time derivatives
so that \(\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)\) and set
\(\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}\) the Laplace
operator on \(\mathbb{R}^d\), except in \S\ref{caC} where \(\Delta\)
represents a random variable.

\hypertarget{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}{%
\subsection{\texorpdfstring{Sufficient conditions for finite mean,
variance and total abundance in the deterministic case
\label{finite}}{Sufficient conditions for finite mean, variance and total abundance in the deterministic case }}\label{sufficient-conditions-for-finite-mean-variance-and-total-abundance-in-the-deterministic-case}}

Recall that \(m(\nu,x)\) is shorthand for \(m\big((K\nu)(x,t),x\big)\).
That is, \(m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}\). Following our
assumptions of the main text, we have that \(m(h,x)\) is differentiable
with respect to both \(x\) and \(h\) and there exists \(r\in\mathbb{R}\)
such that \(m(h,x)\leq r\) across all \(h\geq0\) and \(x\in\mathbb{R}\).
As in the main text, we also assume the initial condition \(u(x)\) is
continuous and integrable in \(x\) and satisfies \begin{equation}
0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
\end{equation} and consider the Cauchy problem
\begin{equation}\label{PDE_SM}
\left\{\begin{matrix}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) & t>0\\
\nu(x,0)=u(x) & t=0.
\end{matrix}\right.
\end{equation}

According to Theorem 2.5.6 of Zheng (2004), if the operator \(F\)
defined by \(\nu(x,t)\to m(\nu,x)\nu(x,t)\) is locally Lipschitz,
corresponding to equation (\ref{local_lipschitz}) of the main text, then
for some maximal \(T>0\) problem (\ref{PDE_SM}) admits a unique local
classical solution \(\nu(x,t)\) for \(t\in[0,T)\). Furthermore, either
\(T=+\infty\) or \(T<+\infty\) and
\(\lim_{t\uparrow T}\int_\mathbb{R}|\nu(x,t)|dx=+\infty\). In this
section we show that our assumption \(m(h,x)\leq r\) for all \(h\geq0\)
and \(x\in\mathbb{R}\) implies \(T=+\infty\). Replacing \(m\) with it's
upper bound \(r\in\mathbb{R}\), equation (\ref{PDE_SM}) reduces to a
simple parabolic equation that can be solved using elementary techniques
(Farlow 1993). In particular, when \(m(\nu,x)\equiv0\) denote the
solution to (\ref{PDE_SM}) by \(\nu_0(x,t)\). Then, denoting
\begin{equation}
\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
\end{equation} we have \begin{equation}
\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
\end{equation} In the more general case, when
\(m(\nu,x)\equiv r\in\mathbb{R}\), equation (\ref{PDE_SM}) has the
solution \(\nu_r(x,t)=e^{rt}\nu_0(x,t)\). Hence, \(\nu_r(x,t)\geq0\) for
all \(x\in\mathbb{R}\) and
\(\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty\) for all \(t\geq0\).
Furthermore, denoting \begin{equation}
\begin{matrix}
N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
\end{matrix}
\end{equation} we have \begin{equation}
\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
\end{equation} \begin{equation}
\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
\end{equation} Hence, \(|\bar x_r(t)|,\sigma_r^2(t)<+\infty\) for all
\(t\geq0\). For the sake of contradiction, suppose there exists
\(x\in\mathbb{R}\) and \(t\geq0\) such that \(\nu(x,t)>\nu_r(x,t)\).
Then \begin{equation}
\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
\end{equation} which implies there exists \(h\geq0\) and
\(x\in\mathbb{R}\) such that \(m(h,x)>r\). But this contradicts our
assumption \(m(h,x)\leq r\) for all \(h\geq0\) and \(x\in\mathbb{R}\).
So we have \(\nu(x,t)\leq\nu_r(x,t)\) for each \(x\in\mathbb{R}\) and
\(t\geq0\). This implies that
\(N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty\), \begin{equation}
0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
\end{equation} and in particular \begin{equation}
0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
\end{equation} for each \(t\geq0\).

\hypertarget{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}{%
\subsection{\texorpdfstring{Equilibrium moments for a deterministic
population experiencing logistic growth and stabilizing selection
\label{equilib}}{Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection }}\label{equilibrium-moments-for-a-deterministic-population-experiencing-logistic-growth-and-stabilizing-selection}}

Here we set out to show, given the initial conditions
\(\nu(\cdot,0)\in L^1(\mathbb{R})\) such that
\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\) and
\(N(0)\in(0,+\infty)\), and given the growth rate
\begin{equation}\label{m_log_stab}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
\end{equation} such that \(\theta\in\mathbb{R}\), \(a,c,\mu>0\) and
\(r>\tfrac{1}{2}\sqrt{\mu a}\), we have the following stable equilibrium
\(N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
\(\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta\) and
\(\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}\).

The mean fitness can be calculated as \begin{equation}
\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t).
\end{equation} Recalling the ODE derived for \(N(t)\), \begin{equation}
\frac{d}{dt}N(t)=\bar m(t)N(t)=(r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-cN(t))N(t),
\end{equation} solving for equilibrium total abundance \(\hat N\)
amounts to setting \(\frac{d}{dt}N(t)=0\) and solving for \(N(t)\).
Ignoring the equilibrium \(N(t)=0\), this reduces to solving
\(\bar m(t)=0\) for \(N(t)\), which returns \begin{equation}
\hat N=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\hat{\bar x})^2-\hat\sigma^2\right)\right).
\end{equation} Unfortunately, deriving ODE for \(\bar x(t)\) and
\(\sigma^2(t)\) leads to expressions involving higher moments and
finding ODE for these higher moments will lead to expressions involving
yet even higher moments. To avoid this infinite regression, we find the
equilibrium abundance density \(\hat\nu(x)\) by solving
\(\frac{\partial}{\partial t}\nu(x,t)=0\) for \(\nu(x,t)\). This implies
the following ordinary differential equation \begin{equation}
\frac{d^2}{dx^2}\hat\nu(x)=\left(\frac{2c}{\mu}\hat N+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\hat\nu(x)
\end{equation} which has the solution \begin{equation}
\hat\nu(x)=\frac{\hat N}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\sqrt\frac{a}{\mu}\frac{(\theta-x)^2}{2}\right).
\end{equation} From this expression we infer \(\hat{\bar x}=\theta\) and
\(\hat\sigma^2=\sqrt{\frac{\mu}{a}}\). Hence
\(\hat N=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)\). To show
that \(N_\infty=\hat N\), \(\bar x_\infty=\hat{\bar x}\) and
\(\sigma^2_\infty=\hat\sigma^2\), we linearize ODE for \(N(t)\),
\(\bar x(t)\) and \(\sigma^2(t)\) using the equilibrium \(\hat\nu(x)\).
In particular, since \(\hat\nu(x)\) is Gaussian, we do not run into the
same issue with higher moments as above. Assuming the equilibrium
\(\hat\nu(x)\) does not change the ODE for \(N(t)\), but the ODE for
\(\bar x(t)\) and \(\sigma^2(t)\) can now be expressed as
\begin{equation}
\frac{d}{dt}\bar x(t)=\sigma^2(t)\left(\frac{\partial\bar m(t)}{\partial\bar x(t)}-\overline{\frac{\partial m(t)}{\partial\bar x(t)}}\right)=a\sigma^2(t)(\theta-\bar x(t)),
\end{equation} \begin{equation}
\frac{d}{dt}\sigma^2(t)=2\sigma^4(t)\left(\frac{\partial\bar m(t)}{\partial\sigma^2(t)}-\overline{\frac{\partial m(t)}{\partial\sigma^2(t)}}\right)+\mu=\mu-a\sigma^4(t),
\end{equation} where \(\hat p(x)=\hat\nu(x)/\hat N\). These expressions
confirm our findings that \(\hat{\bar x}=\theta\) and
\(\hat\sigma^2=\sqrt{\frac{\mu}{a}}\). Furthermore, calculating
\begin{equation}
\frac{\partial}{\partial\sigma^2(t)}\frac{d}{dt}\sigma^2(t)=-2a\sigma^2(t)
\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\)
demonstrates the equilibrium phenotypic variance is stable when
\(a,\mu>0\). Hence, calculating \begin{equation}
\frac{\partial}{\partial\bar x(t)}\frac{d}{dt}\bar x(t)=-a\sigma^2(t)
\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\) and
\(\bar x(t)=\hat{\bar x}\) demonstrates the equilibrium phenotypic mean
is stable when \(a,\mu>0\). Finally, calculating \begin{equation}
\frac{\partial}{\partial N(t)}\frac{d}{dt}N(t)=r-\frac{a}{2}((\theta-\bar x(t))^2+\sigma^2(t))-2cN(t)
\end{equation} and evaluating at \(\sigma^2(t)=\hat\sigma^2\),
\(\bar x(t)=\hat{\bar x}\) and \(N(t)=\hat N\) demonstrates the
equilibrium total abundance is stable when \(a,c,\mu>0\), and
\(r>\frac{1}{2}\sqrt{a\mu}\).

\hypertarget{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}{%
\subsection{\texorpdfstring{The relation between diffusion and
convolution with a Gaussian kernel
\label{diffconvequiv}}{The relation between diffusion and convolution with a Gaussian kernel }}\label{the-relation-between-diffusion-and-convolution-with-a-gaussian-kernel}}

For continuous \(g:\mathbb{R}^d\to\mathbb{R}\), consider the
deterministic Cauchy problem \begin{equation}\label{heateqn}
\left\{\begin{matrix}
\dot f(x,t)=&\Delta f(x,t), & (x,t)\in\mathbb{R}^d\times(0,\infty)\\
f(x,t)=&g(x), & (x,t)\in\mathbb{R}^d\times\{0\}.
\end{matrix}\right.
\end{equation} According to Evans (2010), the fundamental solution of
(\ref{heateqn}) is \begin{equation}
\Phi(x,t)=\frac{1}{(4\pi t)^{d/2}}\exp\left(-\frac{|x|^2}{4t}\right), \ (x,t)\in(0,\infty)\times\mathbb{R}^d,
\end{equation} where \(|x|=\sqrt{\sum_ix_i^2}\). The solution \(f(x,t)\)
of PDE (\ref{heateqn}) is then given by the convolution \begin{equation}
f(x,t)=\int_{\mathbb{R}^d}\Phi(x-y,t)g(y)dy, \ (x,t)\in(0,\infty)\times\mathbb{R}^d.
\end{equation} Hence, by the fundamental theorem of calculus,

\begin{multline}
f(x,t)+\int_t^{t+1}\dot f(x,s)ds=f(x,t+1) \\
=\int_{\mathbb{R}^d}\Phi(x-y,t+1)g(y)dy=\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\Phi(x-y,1)\Phi(y-z,t)g(z)dzdy \\
=\int_{\mathbb{R}^d}\Phi(x-y,1)f(t,y)dy.
\end{multline} In particular, \begin{equation}
f(x,t)+\int_t^{t+1}\Delta f(x,s)ds=\int_{\mathbb{R}^d}\Phi(1,x-y)f(y,t)dy.
\end{equation}

\hypertarget{deterministic-dynamics-of-sigma2t}{%
\subsection{\texorpdfstring{Deterministic dynamics of \(\sigma^2(t)\)
\label{var_deriv}}{Deterministic dynamics of \textbackslash{}sigma\^{}2(t) }}\label{deterministic-dynamics-of-sigma2t}}

Picking up from the main text \S\ref{deterministic},

\begin{multline}
\frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
\end{multline} Applying integration by parts twice yields
\begin{equation}
\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
\end{equation}

\hypertarget{simplifying-covariances-with-fitness-under-the-assumption-of-a-gaussian-density}{%
\subsection{\texorpdfstring{Simplifying covariances with fitness under
the assumption of a Gaussian density
\label{cov2deriv}}{Simplifying covariances with fitness under the assumption of a Gaussian density }}\label{simplifying-covariances-with-fitness-under-the-assumption-of-a-gaussian-density}}

By assuming

\begin{equation}
p(x,t)=\frac{\exp\left(-\frac{(x-\bar x(t))^2}{2\sigma^2(t)}\right)}{\sqrt{2\pi\sigma^2(t)}}
\end{equation}

we have

\begin{multline}
\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial \bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
\end{multline}

and

\begin{multline}
2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)=2\sigma^4\left(\frac{\partial}{\partial\sigma^2}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial\sigma^2}m(\nu,x)dx\right)\\=2\sigma^4\int_\mathbb{R}\frac{(x-\bar x)^2-\sigma^2}{2\sigma^4}m(\nu,x)p(x,t)dx=\int_\mathbb{R}\left((x-\bar x)^2-\sigma^2\right)\big(m(\nu,x)-\bar m\big)p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x)^2,m\Big).
\end{multline}

\hypertarget{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}{%
\subsection{\texorpdfstring{Comparing our treatment of white noise to Da
Prato and Zabczyk (2014)
\label{DPZ14}}{Comparing our treatment of white noise to Da Prato and Zabczyk (2014) }}\label{comparing-our-treatment-of-white-noise-to-da-prato-and-zabczyk-2014}}

Our approach in the main text is inspired by the treatment provided in
\S4.2 of Da Prato and Zabczyk (2014). Here the authors develop a
stochastic integral of operator-valued processes. In particular, they
consider processes indexed by time \(t\geq0\) valued as Hilbert-Schmidt
operators \(\Phi(t)\) and define the norm \begin{equation}
\|\Phi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\mathrm{Tr}[\Phi(s)\Phi^*(s)]ds\right)}, \ t\geq0.
\end{equation} In our case we only consider the so-called multiplication
operators. That is, processes that consist of operators \(\Phi(t)\)
having the form \(\Phi(t)g(x)=\varphi(x,t)g(x)\) such that
\(\varphi(\cdot,t)\in L^2(\mathbb{R})\) a.s. for each \(t\geq0\). In
this case \(\Phi(t)=\Phi^*(t)\) and \begin{equation}
\|\Phi\|_t=\|\varphi\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}\varphi^2(x,s)dxds\right)}, \ t\geq0.
\end{equation} Da Prato and Zabczyk (2014) form the space
\(\mathscr{N}^2_W(0,T)\) of Hilbert-Schmidt operator-valued predictable
processes \(\Phi(t)\) that satisfy \(\|\Phi\|_T<+\infty\) for some
\(T>0\). This corresponds to our more specialized space
\(\mathscr{N}_2\) that consists of \(L^2(\mathbb{R})\) valued processes
\(\varphi(x,t)\) such that \(\|\varphi\|_t<+\infty\) for all \(t\geq0\).
In their treatment, \(W(t)\) plays a similar role to our generalized
process \(\mathbf{W}_t\). For \(\Phi\in\mathscr{N}^2_W(0,T)\), they
denote the stochastic integral for \(t\in[0,T]\) by \(\Phi\cdot W(t)\).
Hence, for \(\Phi(t) g(x)=\varphi(x,t)g(x)\) as above,
\(\mathbf{W}_t(\varphi)=\Phi\cdot W(t)\). The authors then prove the
following:

\begin{quote}
\textbf{Proposition 4.28} \emph{Assume that
\(\Phi_1,\Phi_2\in\mathscr{N}^2_W(0,T)\). Then}

\[\mathbb{E}(\Phi_i\cdot W(t))=0, \ \ \mathbb{E}(\|\Phi_i\cdot W(t)\|^2)<+\infty, \ \ \forall \ t\in[0,T].\]

\textbf{Corollory 4.29} \emph{Under the same assumptions as Proposition
4.28,}

\[\mathbb{C}(\Phi_1\cdot W(t),\Phi_2\cdot W(s))=\mathbb{E}\left(\int_0^{t\wedge s}\mathrm{Tr}[\Phi_2(r)\Phi_1^*(r)]dr\right), \ \ \forall \ t,s\in[0,T].\]
\end{quote}

Simplifying these expressions for the multiplication operators described
above returns equations (\ref{exp_WN_xi}) and (\ref{cov_WN_xi}) of the
main text.

\hypertarget{simulating-the-rescaled-process}{%
\subsection{\texorpdfstring{Simulating the rescaled process
\label{numerical}}{Simulating the rescaled process }}\label{simulating-the-rescaled-process}}

Here we provide a detailed description of the branching random walk and
how we have chosen to rescale it. We focus on the case \begin{equation}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx.
\end{equation}

\hypertarget{description-of-simulation}{%
\subsubsection{\texorpdfstring{Description of simulation
\label{description}}{Description of simulation }}\label{description-of-simulation}}

We begin by describing the branching random walk before introducing our
scheme to rescale it. Our branching random walk follows closely the
description of branching Brownian motion in the main text. However, we
replace exponentially distributed lifetimes with deterministic unit time
steps for easier implementation. Hence, we restrict time to
\(t=0,1,2,\dots,\) and so on. Furthermore, we allow individual fitness
to depend on both trait value and the state of the entire population.
For time \(t\) we write \(\{\xi_1(t),\dots,\xi_{N(t)}(t)\}\) as the set
of trait values across all \(N(t)\) individuals alive in the population.
Since our simulation follows discrete individuals instead of continuous
distributions of trait values, we can write \begin{equation}
\nu(x,t)=\sum_{i=1}^{N(t)}\delta(x-\xi_i(t)),
\end{equation} where \(\delta(x-\xi_i(t))\) denotes the point mass
located at \(\xi_i(t)\). For simplicity we assume perfect heritability.
At each iteration we draw, for each individual, a random number of
offspring from a Negative-Binomial distribution. Recall the
Negative-Binomial distribution models the number of failed Bernoulli
trials that occur before a given number of successful trials. Denoting
\(q\) the probability of success for each trial and \(s\) the number of
successes, the mean and variance is given respectively by
\begin{equation}
\frac{s(1-q)}{q}, \ \frac{s(1-q)}{q^2}.
\end{equation} Then if we require the \(i\)th individual to have mean
number offspring \(\mathscr W(\nu,\xi_i)\) and variance equal to \(V\),
the parameters of the associated Negative-Binomial distribution become
\begin{equation}
q(\nu,\xi_i)=\frac{\mathscr{W}(\nu,\xi_i)}{V}, \ s(\nu,\xi_i)=\frac{\mathscr{W}^2(\nu,\xi_i)}{V-\mathscr{W}(\nu,\xi_i)}.
\end{equation} This imposes the restriction
\(V>\mathscr{W}(\nu,\xi_i)\). For each offpsring produced by the
individual with trait value \(\xi_i(t)\), we assign indepently drawn
trait values normally distributed around \(\xi_i(t)\) with variance
\(\mu\). This summarizes the basic structure of our simulation. To
impose selection and density dependent growth rates, we set
\begin{equation}
\mathscr{W}(\nu,\xi_i)=\exp\left(r-\frac{a}{2}(\theta-\xi_i)^2-c\int_\mathbb{R}\nu(x,t)dx\right),
\end{equation} where the above integral becomes
\(\int_\mathbb{R}\nu(x,t)dx=\sum_{i=1}^{N(t)}1=N(t)\).

\textbf{Rescaling}

To rescale the branching random walk by a positive integer \(n\), we
rescale mutational variance by \(\mu\to \mu/n\), time by \(t\to t/n\)
and the reproductive law by \(V\to V\) and \begin{equation}
\mathscr W(\nu,\xi_i)\to\mathscr W^{(n)}(\nu,\xi_i)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n^2}N(t)\right)=\exp\left(\frac{r}{n}-\frac{a}{2n}(\theta-\xi_i)^2-\frac{c}{n}N^{(n)}(t)\right).
\end{equation} We also replace individual mass with \(\frac{1}{n}\) and
write rescaled abundance as \(N^{(n)}(t)=\frac{1}{n}N(nt)\). When it
exists, we denote by \(N^{(\infty)}(t)\) the limiting process of
\(N^{(n)}(t)\). Then \begin{equation}
\lim_{n\to\infty}n\left(\mathscr W^{(n)}(\nu,\xi_i)-1\right)=r-\frac{a}{2}(\theta-\xi_i)^2-cN^{(\infty)}(t).
\end{equation} Note that, since the limiting fitness function is bounded
above and decreases linearly with respect to \(N^{(\infty)}(t)\), it
satisfies the hypotheses of Champagnat, Ferrière and Méléard (2006). We
have implemented this simulation in the programming language Julia. A
copy can be found at the url:
\[\texttt{\url{https://github.com/bobweek/branching.brownian.motion.and.spde}}\]
For the sake of illustration, we simulated the unscaled process
(\(n=1\)) and the rescaled process with \(n=5\) and \(n=20\) for \(50\)
units of time. Results are shown in Figure \ref{rescaled} of the main
text.

\hypertarget{derivation-of-sde-for-bar-x-and-sigma2}{%
\subsection{\texorpdfstring{Derivation of SDE for \(\bar x\) and
\(\sigma^2\)
\label{SDE_DERIV}}{Derivation of SDE for \textbackslash{}bar x and \textbackslash{}sigma\^{}2 }}\label{derivation-of-sde-for-bar-x-and-sigma2}}

Picking up from \S\ref{stochastic} of the main text, we have
\begin{equation}
\tilde x(t)=\int_\mathbb{R}x\nu(x,t)dx, \ \ \tilde{\tilde\sigma}^2(t)=\int_\mathbb{R}x^2\nu(x,t)dx
\end{equation} and \begin{equation}\label{xtilde}
\tilde x(t)=\tilde x(0)+\int_0^t\int_\mathbb{R}\nu(x,s)m(\nu,x)x+x\sqrt{V\nu(x,s)}\dot W(x,s)dxds,
\end{equation} \begin{equation}\label{stilde}
\tilde{\tilde\sigma}^2(t)=\tilde{\tilde\sigma}^2(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)x^2+\mu\right)+x^2\sqrt{V\nu(x,s)}\dot W(x,s)dxds.
\end{equation}

\hypertarget{derivation-for-trait-mean}{%
\subsubsection{Derivation for trait
mean}\label{derivation-for-trait-mean}}

We make use of the notation

\begin{equation}
\left\{\begin{matrix}
\|N\|_2 = \sqrt{V\int_\mathbb{R}\nu(x,t)dx}= \sqrt{VN} \\ \\
\|\tilde x\|_2 = \sqrt{V\int_\mathbb{R}x^2\nu(x,t)dx} \\ \\
\langle\tilde x,N\rangle = V\int_\mathbb{R}x\nu(x,t)dx=\bar x VN.
\end{matrix}\right.
\end{equation}

Rewriting formula (\ref{xtilde}) as an SDE provides

\begin{equation}
d\tilde x=\left(\overline{xm}N+\frac{\mu}{2}\int_\mathbb{R}x\Delta\nu(x,t)dx\right)dt+\|\tilde x\|_2d\tilde W_2,
\end{equation}

where,

\begin{equation}
d\tilde{ W}_2=d\hat{\mathbf W}_t(\sqrt{V x^2\nu})=\frac{1}{\|\tilde x\|_2}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dxdt.
\end{equation}

Using Itô's quotient rule on \(\bar x=\tilde x/N\), we obtain

\begin{equation}
d\bar x = d\left(\frac{\tilde x}{N}\right)=\frac{\tilde x}{N}\left(\frac{d\tilde x}{\tilde x}-\frac{dN}{N}-\frac{d\tilde x}{\tilde x}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
= \frac{d\tilde x}{N}-\bar x\frac{dN}{N}-\frac{d\tilde x}{N}\frac{dN}{N}+\bar x\Big(\frac{dN}{N}\Big)^2.
\end{equation}

From Table \ref{mult} of the main text
\(d\tilde xdN=\langle\tilde x,N\rangle\) and \(dN^2=\|N\|_2^2\). Hence,

\begin{equation}
d\bar x= \overline{xm}dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\left(\bar m dt+\sqrt\frac{V}{N}dW_1\right)-\frac{\langle\tilde x,N\rangle}{N^2}dt+\bar x\frac{\|N\|_2^2}{N^2}dt
\end{equation}

\begin{equation*}
= (\overline{xm}-\bar x\bar m)dt+\frac{\|\tilde x\|_2}{N}d\tilde W_2
-\bar x\sqrt{\frac{V}{N}}d W_1-V\frac{\bar x}{N}dt+V\frac{\bar x}{N}dt
\end{equation*}

\begin{equation*}
=\mathrm{Cov}_t(x,m)
+\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1.
\end{equation*}

Note that

\begin{equation}
\frac{\|\tilde x\|_2}{N}d\tilde W_2-\bar x\sqrt{\frac{V}{N}}d W_1=\frac{1}{N}\int_\mathbb{R}x\sqrt{V\nu(x,t)}\dot W(x,t)dx-\frac{\bar x}{N}\int_\mathbb{R}\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation}

\begin{equation*}
=\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx
\end{equation*}

and

\begin{equation}
\mathbb{V}\left(\int_\mathbb{R}\frac{x-\bar x}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx\right)=\frac{V}{N}\int_\mathbb{R}(x-\bar x)^2p(x,t)dx=V\frac{\sigma^2}{N}.
\end{equation}

Hence, by setting

\begin{equation}
d W_2=\frac{\int_\mathbb{R}\frac{(x-\bar x)}{N}\sqrt{V\nu(x,t)}\dot W(x,t)dx}{\sqrt{V\sigma^2/N}}
\end{equation}

we can write

\begin{equation}
d\bar x=\mathrm{Cov}_t(x,m)dt
+\sqrt{V\frac{\sigma^2}{N}}d W_2.
\end{equation}

\hypertarget{derivation-for-trait-variance}{%
\subsubsection{Derivation for trait
variance}\label{derivation-for-trait-variance}}

We make use of the notation

\begin{equation}
\left\{\begin{matrix}
\|\tilde{\tilde\sigma}^2\|_2 = \sqrt{V\int_\mathbb{R}x^4\nu(x,t)dx} \\ \\
\langle\tilde{\tilde\sigma}^2,N\rangle=V\int_\mathbb{R}x^2\nu(x,t)dx=\overline{x^2}VN.
\end{matrix}\right.
\end{equation}

Applying formula (\ref{stilde}) provides

\begin{equation}
d\tilde{\tilde\sigma}^2=\left(\overline{x^2m}N+\mu N\right)dt+\|\tilde{\tilde\sigma}^2\|_2d\tilde W_3
\end{equation}

\begin{equation}
d\tilde{W}_3=d\hat{\mathbf W}_t(\sqrt{V x^4\nu})
=\frac{1}{\|\tilde{\tilde\sigma}^2\|_2}\int_\mathbb{R}x^2\sqrt{V\nu(x,t)}\dot W(x,t)dx.
\end{equation}

Using Itô's quotient rule on
\(\overline{x^2}=\tilde{\tilde\sigma}^2/N\), we obtain

\begin{equation}
d\overline{x^2} = d\left(\frac{\tilde{\tilde\sigma}^2}{N}\right)=\frac{\tilde{\tilde\sigma}^2}{N}\left(\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}-\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{\tilde{\tilde\sigma}^2}\frac{dN}{N}+\Big(\frac{dN}{N}\Big)^2\right)
=\frac{d\tilde{\tilde\sigma}^2}{N}-\overline{x^2}\frac{dN}{N}-\frac{d\tilde{\tilde\sigma}^2}{N}\frac{dN}{N}+\overline{x^2}\left(\frac{dN}{N}\right)^2.
\end{equation}

Table \ref{mult} of the main text implies
\(d\tilde W_3d W_1=\langle\tilde{\tilde\sigma}^2,N\rangle\) and hence

>>>>

\usepackage{xr}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage[left]{lineno}
\usepackage{amsmath}
\linenumbers
\externaldocument{SM}
\usepackage{float}
\usepackage{flafter}

@@
